% !TeX program = lualatex
% !TeX encoding = utf-8

\documentclass[master.tex]{subfiles}
% \documentclass[11pt]{report}
% \input{./preamble.tex}

\begin{document}
\section{Real numbers as Dedekind cuts}
Now we can construct $\R$ in a way that shows it to be both a complete, ordered field and a subset of $2^{\Q}$ by using \emph{Dedekind cuts}.

\begin{definition}{Dedekind cut}{dedekind-cut}
    A \textbf{Dedekind cut} is a set $A \subseteq \Q$ such that:
    \begin{itemize}
        \item $\forall a \in A \quad b \leq a \implies b \in A$.
        \item $\forall a \in A \quad \exists a' > a$ such that $a' \in A$.
    \end{itemize}
\end{definition}

The first condition says a Dedekind cut is open downward, i.e.\ it continues to $-\infty$.
The second condition says it is open upward, i.e.\ it has no maximal element in the cut.
(Note that this does not necessarily mean it goes to $\infty$).

\begin{example}{}{r-dedekind}
    Let $R := \set{D \subseteq \Q : D \text{ is Dedekind}}$ and define $\phi : \Q \rightarrow R$ by $q \mapsto \set{a : a < q}$.
    Then $R$ is Dedekind.
\end{example}

For a visual interpretation of Example \ref{ex:r-dedekind} (and setting aside for a moment that we do not know what ``real numbers'' are yet), imagine cutting the real line at every rational $q$.
Every rational to the left of $q$ is in the cut $D$, and no matter which $a \in D$ we pick, we can always find another rational to the right of $a$ that is arbitrarily less than $q$ and therefore in $D$.

\begin{example}{}{sqrt-2-dedekind}
    $A_{\sqrt{2}} := \set{x \in \Q : x^2 < 2 \text{ or } x \leq 0}$ is Dedekind.
\end{example}

Try showing that these example sets satisfy Definition \ref{def:dedekind-cut}.

Since $\Q$ is totally ordered, Example \ref{ex:r-dedekind} suggests that we can also order Dedekind cuts.
Indeed, this shows us that that set $R$ of Dedekind cuts is complete.

\begin{definition}{}{}
    Let $D_1, D_2 \in R$.
    If $D_1 \subseteq D_2$, then we say $D_1 \leq D_2$.
\end{definition}

\begin{lemma}{}{set-of-dedekind-complete}
    $R$ is complete.
    \hr{}
    \begin{proof}
        Let $S \subset R$ be a nonempty set of Dedekind cuts and let $D \in R$ be an upper bound for $S$.
        Then for every $s \in S$, by definition $s \leq D$ and $s \subseteq D$.

        Now define $\sup{S} := \bigcup_{s \in S} s$.
        This union is trivially contained in $D$, and we can also confirm it is in $R$ by showing it verifies the conditions of Definition \ref{def:dedekind-cut}.
        Thus, $\sup{S} \leq D$.

        To show that $\sup{S}$ is indeed a supremum, we must show that it is the least upper bound for $S$.
        Let $D' \in R$ be another upper bound.
        Then for every $s \in S$ we have $s \subseteq D'$, which implies $\sup{S} = \cup_{s \in S} s \subseteq D'$ and therefore $\sup{S} \leq D'$.
    \end{proof}
\end{lemma}

However, is $R$ the real numbers that we are familiar with?
Not quite, because this set is too big!
Note that $\Q$ itself is a Dedekind cut in $R$ that would correspond to $\infty$, which we know is not a number.
Hence, we refine our definition of $R$ to give a proper definition of the \emph{real numbers}.

\begin{definition}{Real numbers}{}
    The \textbf{real numbers} $\R$ are the set of proper Dedekind cuts
    \[
        \R := \set{D \subset \Q : D \neq \Q \text{ is Dedekind}}
    .\]
    If $D \neq \Q$, then take any $q \notin D$.
    No $q' > q$ can be in $D$, so we can equivalently define $\R$ by
    \[
       \R := \set{D \in R : \exists q \in \Q \text{ such that } D < \phi(q)}
    ,\]
    where $\phi : \Q \rightarrow \R$ is analagously defined as in Example \ref{ex:r-dedekind}.
\end{definition}

\begin{remark}
    $\R$ is complete.
    \hr{}
    \begin{proof}
        If $D \in \R$, then by definition there exists $q_0$ such that $D < \phi(q_0)$.
        This implies
        \[
            \sup{D} < \phi(q_0) \implies \sup{D} \neq \infty \implies \sup{D} \in \R
        .\]
        The rest of the proof is the same as for Lemma \ref{lem:set-of-dedekind-complete}.
    \end{proof}
\end{remark}

\hr{}

We now consider the field nature of $\R$.
Define the \textbf{positive reals} $\R^{+}$ as the set
\[
    \R^{+} := \set{D \in \R : (-\infty, 0] \subseteq D}
,\]
and define addition and multiplication in it by
\begin{align*}
    D_1 + D_2 &= \bigcup \set{d_1 + d_2 : D_1^{+}, d_2 \in D_2^{+}} \\
    D_1 D_2 &= \bigcup \set{d_1 d_2 : d_1 \in D_1^{+}, d_2 \in D_2^{+}}.
\end{align*}
Since negative numbers complicate multiplication, we can assume that $D \in \R^{+}$ is a subset of $\Q^{+}$ and ``starts'' at $0$ rather than $-\infty$.
More formally, we can view this as $\R^{+} = \set{D \subset \Q^{+} : D \cup (-\infty, 0] \text{ is Dedekind}}$.


With these operations, $\R^{+}$ satisfies the field axioms with associative, commutative, and distributive properties (verify these for practice!):
\begin{align*}
    D_1 + (D_2 + D_3) &= (D_1 + D_2) + D_3 \\
    D_1 (D_2 + D_3) &= D_1 D_2 + D_1 D_3 \\
    &\vdotswithin{=}
\end{align*}
Indeed, $\R^{+}$ is a commutative and cancellative semigroup under $+$.
Our final answer to what the real numbers are is that $\R$ is the Grothendieck group for $\R^{+}$:
\[
    \R = G(\R^{+}, +)
.\]

\begin{lemma}{}{}
    $\R$ is totally ordered.
    \hr{}
    \begin{proof}
        Let $D_1, D_2 \in \R$.
        There are two cases:
        \begin{enumerate}
            \item[] Case 1: $D_1 \subseteq D_2$.
            Then by definition $D_1 \leq D_2$, and we are done.

            \item[] Case 2: $D_1 \not\subseteq D_2$.
            Then there exists $d_1 \in D_1$ that is not in $D_2$.
            This means for every $d_2 \in D_2$ we have $d_2 < d_1$.
            By definition, this implies $D_2 \subseteq D_1$ and hence $D_2 \leq D_1$.
        \end{enumerate}
    \end{proof}
\end{lemma}

    \subsection*{Exercises}
    \begin{enumerate}[font=\bfseries]
        \item Let $\phi : \Q^{+} \rightarrow \R^{+}$ (where $\R^{+}$ is a set of Dedekind cuts) be the function defined by
        \[
            \phi(q) := \set{a \in \Q : a < q}
        .\]
        \begin{enumerate}[font=\bfseries]
            \item Prove that $\phi$ is injective and order-preserving, i.e.\ show that if $q < q'$, then $\phi(q) < \phi(q')$.
            \item Prove that $\phi$ respects both field operations:
            \[
                \phi(qr) = \phi(q) \phi(r) \text{ and } \phi(q + r) = \phi(q) + \phi(r)
            .\] 
        \end{enumerate}

        \item Let $A_{\sqrt{2}}$ be the Dedekind cut from Example \ref{ex:sqrt-2-dedekind}.
        Prove that $\sup{A_{\sqrt{2}}} = \sqrt{2}$.
    \end{enumerate}
\end{document}
